---
title: "test3"
output: html_document
date: "2023-02-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## q1

Given IQ scores are approximately normally distributed with a mean of 100 and standard
deviation of 15, the proportion of people with IQs above 130 is:

```{r cars}
pnorm(130,mean=100,sd=15,lower.tail=FALSE)


```
## q2

In 1923, Babe Ruth had 522 at bats with 205 hits. Assuming that the binomial distribution can be appropriately applied, find the expected number of hits in 529 at bats.

```{r cars}

prob <- 205/522
prob * 529


```
## q3

(7) Use the information provided to determine a 95% confidence interval for the population variance. 

A researcher was interested in the variability in service time (in hours) spent by mechanics fixing the same automotive problem.  A random sample was taken resulting in a sample of size 20 from a substantial file of reported experience.  The summary statistics are as follows:  n = 20, sample mean = 13.8 hours, sample standard deviation = 3.9 hours.  Assume service time follows a normal distribution.  Round to two decimal places.

```{r pressure, echo=FALSE}


q3_n <- 20
q3_mean <- 13.8
q3_std <- 3.9

q3_ci <- .95
q3_alpha <- 1 - q3_ci

q3_alpha <- qchisq(0.025,df=19)
q3_minusalpha <- qchisq(0.975,df=19)
q3_alpha
q3_minusalpha

qchisq(c(.025,.975),df=19, lower.tail=FALSE)
```

## q7
(6) Use the information to find the minimum sample size required to estimate an unknown population mean µ.  Margin of error: $110, confidence level: 95%, σ = $500.

```{r pressure, echo=FALSE}


q7_err <- 110
q7_ci <- .95
q7_sig <- 500

#e = standard error * zscore 
#500/sqrt(n) * 1.96 = 110
#500/sqrt(n) = 110/1.96
#500 = 56.12244 * sqrt(n)
#sqrt(n) = 500/56.12244
zstar <- qnorm(.975) 
sigma <- 500 
E <- 110 
(zstar^2)*(sigma^2/E^2)


```

## q8

Suppose the respective probabilities are 0.7, 0.2, and 0.1 that a person applying for a driver’s
license in Saskatchewan will require 1, 2, or 3 attempts in order to obtain a license. Let X
be a random variable, the number of attempts in order to obtain a license. Find the mean
(expected value) of X.

```{s}
#define values
vals <- c(1,2,3)

#define probabilities
probs <- c(.7,.2,.1)
sum(vals*probs)
```

## q9

If you wanted to estimate the true percentage of all voters in Canada who are in favor of
abolition of the senate, and if you wanted your maximum error of estimate to be 6% with
a confidence level of 95%, what would the required sample size be?

```{r pressure}

zscore <- 1.96
zsq <- zscore^2
q9p <- .5
esquared <- .0036

#me = 1.96 * sqrt(p(1-p)/n)
#.06 = 1.96 * sqrt(.25/n)
# .06/1.96 = sqrt(.25/n)
# (.06/1.96)^2 = .25/n
# n = .25/((.06/1.96)^2)
.25/((.06/1.96)^2)
```

## q10

(10) Assume normality and use the information given to find the p-value.  Based on the p-value estimated, determine if the null hypothesis should be rejected at a 0.1 significance level.  Select the correct answer if the test statistic in a two-tailed test is z= -1.63.  Note: Remember that it is a 2-sided test.

```{r pressure}

2*pnorm(q=-1.63)
pnorm(q=-1.63, lower.tail=TRUE)
```